---
title: "p8105_hw2_geg2145"
author: "Gustavo Garcia-Franceschini"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1

First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

```{r}
df_pols = read_csv(file = "./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon, 
           sep= "-", 
           into = c("year", "month", "day"),
           convert = T) %>%
  mutate(month = month.name[month],
         president = ifelse(prez_dem == 1, "dem", "gop")) %>%
  select(-c("day", "prez_dem", "prez_gop"))
```

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r}
df_snp = read_csv(file = "./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(col = date, 
           sep= "/", 
           into = c("month", "day", "year"),
           convert = T) %>%
  mutate(month = month.name[month],
         year = ifelse(year < 50, year + 2000, year + 1900)) %>%
  arrange(year, month) %>%
  relocate(year, month)
```

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r}
df_unemp = read_csv(file = "./data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(jan:dec,
               names_to = "month",
               values_to = "unemployment_rate") %>%
  mutate(month = case_when(
    month == "jan" ~ "January", 
    month == "feb" ~ "February", 
    month == "mar" ~ "March", 
    month == "apr" ~ "April", 
    month == "may" ~ "May", 
    month == "jun" ~ "June", 
    month == "jul" ~ "July", 
    month == "aug" ~ "August", 
    month == "sep" ~ "September", 
    month == "oct" ~ "October", 
    month == "nov" ~ "November", 
    month == "dec" ~ "December"
  )) %>%
  arrange(year, month)
```

Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r}
df_p1 = df_pols %>%
  left_join(., df_snp, by = c("year", "month")) %>%
  left_join(., df_unemp, by = c("year", "month"))
```

Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

**drvrjvnrv**

# Problem 2


