---
title: "p8105_hw2_geg2145"
author: "Gustavo Garcia-Franceschini"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Problem 1

First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

```{r p1_pols}
df_pols = read_csv(file = "./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon, 
           sep= "-", 
           into = c("year", "month", "day"),
           convert = T) %>%
  mutate(month = month.name[month],
         president = ifelse(prez_dem == 1, "dem", "gop")) %>%
  select(-c("day", "prez_dem", "prez_gop"))
```

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r p1_snp}
df_snp = read_csv(file = "./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(col = date, 
           sep= "/", 
           into = c("month", "day", "year"),
           convert = T) %>%
  mutate(month = month.name[month],
         year = ifelse(year < 50, year + 2000, year + 1900)) %>%
  arrange(year, month) %>%
  relocate(year, month)
```

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r p1_unemp}
df_unemp = read_csv(file = "./data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(jan:dec,
               names_to = "month",
               values_to = "unemployment_rate") %>%
  mutate(month = case_when(
    month == "jan" ~ "January", 
    month == "feb" ~ "February", 
    month == "mar" ~ "March", 
    month == "apr" ~ "April", 
    month == "may" ~ "May", 
    month == "jun" ~ "June", 
    month == "jul" ~ "July", 
    month == "aug" ~ "August", 
    month == "sep" ~ "September", 
    month == "oct" ~ "October", 
    month == "nov" ~ "November", 
    month == "dec" ~ "December"
  )) %>%
  arrange(year, month)
```

Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r p1_final_df}
df_p1 = df_pols %>%
  left_join(., df_snp, by = c("year", "month")) %>%
  left_join(., df_unemp, by = c("year", "month"))
```

Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

**The `pols` dataset had information on the number of politicians that were either Republican or Democrat at specific dates. In particular, we were interested in whether the president was Republican or Democrat (the `president` variable). The `snp` data contained the `close` variable, which is the closing value of the S&P stock index for specific dates. Likewise, the `unemployment` dataset had the `unemployment rate` in the US for a specific date. The resulting dataset has these variables, with the date (`year`-`month` combinations) being useful for merging. The data spans from `r min(pull(df_p1, year))` to `r max(pull(df_p1, year))`, has `r nrow(df_p1)` rows and `r ncol(df_p1)` columns.**

# Problem 2


